Precise spiking motifs in neurobiological and neuromorphic data

### 2022-11-16 manuscript.tex
figures/izhikevich-repro.svg

\citep{hussain_deltron_2012} implemented a delay-learning algorithm on an analog chip. Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}. In addition to delay learning, a group, in the University of West Sydney, developed a neurmorphic implementation of multiple synaptic plasticity learning rules~\citep{wang_neuromorphic_2015}. They show that STDP and Spike Timing Dependant Delay Plasticity rules could be implemented in both a digital and an analog chip. From the network parameters and the physical limitations to store it on-chip, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network. The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citep{wang_fpga_2014}. %This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance. 

@@

A recent work makes an implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}. They highlight the fact that using delays can be expensive because incoming spikes have to be stored in blocks with a temporal dimension equilavent to the maximal delay. Note that this problem is also due the algorithm used in this paper and that the analog chips must not suffer from this type of limitations.

@@

A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output. However, we have seen that the neurobiological system uses processing loops within hierarchical graphs. In general, these systems are bidirectionally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}. The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies.
%
\subsection{Perspectives}
%
The limits that we have presented can be treated individually in each model, as evidenced by individual efforts which try to overcome them. However, to propose a real breakthrough, we believe that future venues should provide with a unified, interdisciplinary approach, with applications to real-world, ecological scenarios and with open and reproducible methods.

>>>>

\citep{hussain_deltron_2012} implemented a delay-learning algorithm on an analog chip. Online learning on neuromorphic chips is still a challenge today and for this work, only the detection of spiking motifs was done on the analog architecture, training was performed digitally and based on the tempotron learning rule~\citep{gutig_tempotron_2006}. In addition to delay learning, a group at the University of West Sydney developed a neuromorphic implementation of multiple synaptic plasticity learning rules~\citep{wang_neuromorphic_2015}. They show that STDP and Spike Timing Dependent Delay Plasticity rules could be implemented in both a digital and an analog chip. From the network parameters and the physical limitations to store it on-chip, they prove that the digital implementation is way easier to scale up and that an external memory would be needed for a larger network. The same group presented a FPGA hardware implementation of polychronous networks in which propagation delays are learned in a supervised manner, based on the expected firing time of the post-synaptic neuron~\citep{wang_fpga_2014}. %This paper proposes a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance. 

@@

A recent work makes an implementation of a sparse Vector Symbol Architecture binding operation on the Loihi chip, delay lines and coincidence detection are used to compute the binding operation~\citep{renner_sparse_2022}. They highlight the fact that using delays can be expensive because incoming spikes have to be stored in blocks with a temporal dimension equivalent to the maximal delay. Note that this problem is also due the algorithm used in this paper and that the analog chips must not suffer from this type of limitations.

@@

A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output. However, we have seen that the neurobiological system uses processing loops within hierarchical graphs. In general, these systems are bidirectionally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}. The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies. Moreover,  these processes have to be distinguished from judgements on timing such as temporal order processing (judging when one event happens relative to another) or duration estimation (measuring how long an event lasts).
%
\subsection{Perspectives}
%
The limits that we have presented can be treated individually in each model, as evidenced by individual efforts which try to overcome them. However, to propose a real breakthrough, we believe that future venues should provide with a unified, interdisciplinary approach, with applications to real-world, ecological scenarios and with open and reproducible methods.
